\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{hyperref}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Blockchain solutions for incentivization of market-linked upskilling in labor markets}

\author{\IEEEauthorblockN{1\textsuperscript{st} Thomas Scholtz}
	\IEEEauthorblockA{\textit{Department of Computer Science} \\
		\textit{University of Stellenbosch}\\
		Stellenbosch, South Africa \\
		21681147@sun.ac.za}
	\and
	\IEEEauthorblockN{2\textsuperscript{nd} Brink Van Der Merwe}
	\IEEEauthorblockA{\textit{Department of Computer Science} \\
		\textit{University of Stellenbosch}\\
		Stellenbosch, South Africa \\
		abvdm@sun.ac.za}

}

\maketitle

\begin{abstract}
	Blockchain technology has proved to be useful for basic applications such as peer-to-peer electronic cash systems i.e., Bitcoin~\cite{nakamoto2008bitcoin}.  Since this milestone, the technology has evolved to support advanced applications which require more elaborate feature sets. Ethereum~\cite{buterin2014ethereum} is an example of one such form of blockchain technology that allows public deployment of capture-resistant, permissionless contracts, resulting in a platform on which interested parties may attempt to design systems that better address systemic social challenges.

	This paper proposes blockchain solutions for incentivization of market-linked upskilling in South African labor markets with intent to address persistent unemployment in Stellenbosch, South Africa - a problem that is hypothesized to be due in part to the cold-start problem of underprivileged individuals in a dual-speed economy.

	The incentive scheme comprises three blockchain solutions: incentive tokens disbursed to learners on completion of course milestones, verified credentials assigned to learners for completion of courses, and HyperCerts awarded to educators for the effective upskilling of learners via courses.
	The solution(s) is investigated in distinct parts and as a whole by: architecting the components of the solution; simulating the resulting scheme - the sum of the solutions; and surveying relevant stakeholders on the perceived efficacy of such a solution with respect to resolving structural unemployment in the town of Stellenbosch.

\end{abstract}

\begin{IEEEkeywords}
	blockchain, tokenomics, HyperCert, Ethereum, incentivisation, unemployment
\end{IEEEkeywords}

\section{Introduction}
Blockchain technology has proved valuable for applications ranging from peer-to-peer electronic cash systems such as Bitcoin~\cite{nakamoto2008bitcoin} to general-purpose computing platforms such as Ethereum~\cite{buterin2014ethereum}. Ethereum offers a Turing-complete scripting environment that enables developers to encode arbitrary rules for ownership, transaction formats, and state transitions. This potential has given rise to a large and growing class of decentralized applications (DApps), with recent surveys cataloguing thousands of deployed DApps across domains including finance, gaming, governance, and identity~\cite{zheng2023dapp}. The emergence of a composable platform with fundamental properties of decentralization, persistence, and auditability, poses a strong case for the design of systems that address pressing systemic social challenges.

Realising ambitious application-layer systems on Ethereum has historically been constrained by the base layer's limited throughput and high transaction costs. A substantial body of recent work addresses these constraints through Layer 2 (L2) scaling architectures including state channels, sidechains, and, most prominently, optimistic and zero-knowledge rollups~\cite{sguanci2021layer2,thibault2022rollups}. Rollup-based L2s now offer orders-of-magnitude higher throughput while inheriting the security guarantees of the underlying L1, rendering them sufficiently mature to reliably host production-scale social and economic applications. It is recognized that there are still pressing technical challenges to be overcome by the Ethereum network (cite); however, the current utility offered by the network is sufficient for the purposes of this research. The maturity of the network is confirmed by an accelerated interest in blockchain for social good, among other domains. Sansone et al ~\cite{sansone2023blockchain} identify reliability, transparency, decentralization, and accessibility as the properties through which blockchain can support social innovation, while earlier work documents proof-of-concept deployments spanning humanitarian aid, supply-chain transparency, digital identity, and philanthropy~\cite{marino2018blockchain}. Much of this work nevertheless remains exploratory, with limited engagement with specific, deeply rooted socio-economic problems in a particular context.

South Africa presents one such context. Youth unemployment has persisted above forty-five percent for more than a decade and, in Stellenbosch in particular, is driven in significant part by a dual-speed economy in which underprivileged entrants face a structural cold-start problem: they cannot acquire market-relevant experience without initial access to the labour market, and cannot access the labour market without prior experience or at least highly relevant skills (mostly gated by experience in labour markets) ~\cite{bhorat2016youth,graham2015youth,francis2019inequality}.
Supply-side interventions targeting general employment skills alone have been shown to be insufficient; they must be coupled with demand-side incentives that connect learners, educators, and employers ~\cite{mckenzie2017almp,bhorat2016youth}. It is reasonable to suggest that sophisticated incentives for employers to inform the upskilling of the unemployed may create the appropriate dynamic that is needed for impactful unemployment intervention. There are three benefits in such a dynamic that are immediately apparent: the unemployed learn skills that are highly relevant for employment; the employers are able to 'pre-train' interested individuals from the unemployed population for potential employment opportunities at the employer; the employers are rewarded with financial incentive. However, the cold-start problem, where unemployed individuals lack the means to provide for themselves in the short-term, still remains. The proposed solution to this is the disbursement of incentive tokens on completion of course milestones, which can be used to purchase basic goods at a network of pre-approved vendors, who exchange the tokens for ZAR stablecoin, and ZAR stablecoin for ZAR. The ZAR stablecoins are either funded by government municipalities, or by Corporate Social Responsibility programs of the employer-educators.

While the merit of this hypothesis is subject to further social studies contextualised to Stellenbosch, the focus of this paper is on the technical feasibility of such a system, making the large assumption that the hypothesis is true in the context of this research.

This paper recognises Ethereum's support for designing application-layer systems and the evident/implicit public demand for such systems in the South African labour market context, and therefore proposes a blockchain-based incentivization scheme for market-linked upskilling in the South African labour market, composed of three interrelated solutions: fungible incentive tokens disbursed to learners on the completion of course milestones; non-transferable, or soulbound, credential tokens assigned to learners for course completion, in accordance to the decentralized society proposal of Weyl et al.~\cite{weyl2022decentralized}; and HyperCerts~\cite{dalrymple2022hypercerts} awarded to educators for the effective upskilling of learners. The combined ecosystem is designed using frameworks that treat incentive design, governance, and tokenomics as interdependent pillars of a token economy~\cite{oliveira2018token,kivilo2026designing,sadykhov2023detect}. The remainder of the paper architects each component, simulates the resulting system to study its tokenomics and tune emerging parameters, and surveys stakeholders on the perceived efficacy of such a scheme in addressing structural unemployment.

\section{Background and Related Work}
\textit{Scope: blockchain-for-social-good, token-economy frameworks~\cite{oliveira2018token,kivilo2026designing,sadykhov2023detect}, credentialing and decentralized society~\cite{weyl2022decentralized}, and the HyperCerts primitive~\cite{dalrymple2022hypercerts}. Glosses tokenomics and HyperCert terminology used in subsequent sections so later material reads cleanly.}

\section{System Architecture}
\textit{The design contribution. Stakeholder ontology and requirements open the section as the contract against which the architecture is evaluated; the remaining subsections specify each component and their on-chain integration.}

\subsection{Stakeholder Ontology and Requirements}
\textit{Stakeholder ontology --- learners, educators, employers, vendors, municipalities, and CSR funders --- followed by a formalisation of the cold-start dynamic and the derived functional and non-functional requirements the system must satisfy.}

\subsection{Overview and Actor/Flow Diagram}
\textit{High-level system diagram showing the three blockchain components, their on-chain interactions, and the off-chain redemption and funding flows linking actors.}

\subsection{Incentive Tokens}
\textit{Fungible incentive tokens disbursed to learners on milestone completion: ERC-20 vs.\ ERC-1155 selection, disbursement mechanism, and the vendor-mediated ZAR-stablecoin redemption loop funded by municipalities or CSR programs.}

\subsection{Credential Tokens}
\textit{Non-transferable (soulbound) tokens issued on course completion. ERC selection, revocability semantics, and alignment with the decentralized society proposal of Weyl et al.~\cite{weyl2022decentralized}.}

\subsection{Educator HyperCerts}
\textit{HyperCerts~\cite{dalrymple2022hypercerts} awarded to educators in proportion to verified upskilling outcomes; impact attribution model and how educators surface for ex-post funding by employers or CSR donors.}

\subsection{Integration and On-chain Interactions}
\textit{End-to-end interactions across the three components: how a milestone completion event triggers token disbursement, credential issuance, and HyperCert attribution, including trust boundaries and required oracles.}

\section{Simulation}
\textit{Tokenomics model, parameter choices, scenarios, and sensitivity results. Translates the requirements of Section III into measurable outcomes (token velocity, redemption pressure, educator-reward distribution) and tunes parameters against them.}

\section{Stakeholder Survey}
\textit{Instrument design, sample, and results from surveying Stellenbosch stakeholders on the perceived efficacy of the proposed scheme. Methods-light, findings-forward.}

\section{Discussion}

\subsection{Threats to Validity}
\textit{Internal, external, and construct threats across the simulation and survey strands.}

\subsection{Deployment and Regulatory Considerations}
\textit{South African regulatory implications for the ZAR-stablecoin leg: SARB position on stablecoins and crypto assets, FSCA designation as a financial product, and KYC/AML obligations on vendors and redemption rails.}

\subsection{Limitations}
\textit{Bounds on the technical-feasibility claim, dependence on hypothesised social dynamics, and known gaps the present design does not resolve.}

\subsection{Future Work}
\textit{Pilot deployment, integration with existing TVET and SETA programs, governance over CSR funding pools, and extensions to broader public-goods funding via HyperCerts.}

\section{Conclusion}
\textit{Recap of the three-component design, summary of simulation and survey findings, and a closing statement on technical feasibility under the stated hypothesis.}

\begin{thebibliography}{00}
	\bibitem{nakamoto2008bitcoin} S. Nakamoto, ``Bitcoin: A peer-to-peer electronic cash system,'' 2008. [Online]. Available: https://bitcoin.org/bitcoin.pdf

	\bibitem{buterin2014ethereum} V. Buterin, ``A next-generation smart contract and decentralized application platform,'' Ethereum White Paper, 2014. [Online]. Available: https://ethereum.org/en/whitepaper/

	\bibitem{zheng2023dapp} P. Zheng, Z. Jiang, J. Wu, and Z. Zheng, ``Blockchain-based decentralized application: A survey,'' \emph{IEEE Open Journal of the Computer Society}, vol.~4, pp.~121--133, 2023, doi: 10.1109/OJCS.2023.3251854.

	\bibitem{sguanci2021layer2} C. Sguanci, R. Spatafora, and A. M. Vergani, ``Layer 2 blockchain scaling: A survey,'' 2021, arXiv:2107.10881. [Online]. Available: https://arxiv.org/abs/2107.10881

	\bibitem{thibault2022rollups} L. T. Thibault, T. Sarry, and A. S. Hafid, ``Blockchain scaling using rollups: A comprehensive survey,'' \emph{IEEE Access}, vol.~10, pp.~93039--93054, 2022, doi: 10.1109/ACCESS.2022.3200051.

	\bibitem{sansone2023blockchain} G. Sansone, D. Leone, and F. Schiavone, ``Blockchain for social good and stakeholder engagement: Evidence from a case study,'' \emph{Corporate Social Responsibility and Environmental Management}, vol.~30, no.~5, pp.~2182--2192, 2023, doi: 10.1002/csr.2477.

	\bibitem{marino2018blockchain} F. Marino and A. Juels, ``Blockchain for social good,'' in \emph{Proc. 4th EAI Int. Conf. Smart Objects and Technologies for Social Good (GoodTechs)}, Bologna, Italy, 2018, pp.~166--168, doi: 10.1145/3284869.3284881.

	\bibitem{bhorat2016youth} H. Bhorat, A. Cassim, and D. Tseng, ``The state of youth unemployment in South Africa,'' Brookings Institution, Washington, DC, Policy Brief, 2016.

	\bibitem{graham2015youth} L. Graham and C. Mlatsheni, ``Youth unemployment in South Africa: Understanding the challenge and working on solutions,'' in \emph{South African Child Gauge 2015}, A. De Lannoy, S. Swartz, L. Lake, and C. Smith, Eds. Cape Town, South Africa: Children's Institute, University of Cape Town, 2015, pp.~51--59.

	\bibitem{francis2019inequality} D. Francis and E. Webster, ``Poverty and inequality in South Africa: Critical reflections,'' \emph{Development Southern Africa}, vol.~36, no.~6, pp.~788--802, 2019, doi: 10.1080/0376835X.2019.1666703.

	\bibitem{weyl2022decentralized} E. G. Weyl, P. Ohlhaver, and V. Buterin, ``Decentralized society: Finding Web3's soul,'' SSRN, Working Paper 4105763, May 2022. [Online]. Available: https://papers.ssrn.com/sol3/papers.cfm?abstract\_id=4105763

	\bibitem{dalrymple2022hypercerts} D. A. Dalrymple, H. Holtebrinck, T. Kuhn, J. Miller, R. L. Ocampo, and P. Schmidt, ``Hypercerts: A new primitive for public goods funding,'' Protocol Labs, Whitepaper, 2022. [Online]. Available: https://www.protocol.ai/blog/hypercert-new-primitive/

	\bibitem{oliveira2018token} L. Oliveira, L. Zavolokina, I. Bauer, and G. Schwabe, ``To token or not to token: Tools for understanding blockchain tokens,'' in \emph{Proc. 39th Int. Conf. Information Systems (ICIS)}, San Francisco, CA, USA, Dec. 2018.

	\bibitem{kivilo2026designing} S. Kivilo, A. Norta, M. Hattingh, S. Avanzo, and L. Pennella, ``Designing a token economy: Incentives, governance, and tokenomics,'' 2026, arXiv:2602.09608. [Online]. Available: https://arxiv.org/abs/2602.09608

	\bibitem{sadykhov2023detect} A. Sadykhov, G. Goodell, D. Dhinakaran, P. Horshalom, and P. Treleaven, ``Decentralized token economy theory (DeTEcT): Token pricing, stability and governance for token economies,'' \emph{Frontiers in Blockchain}, vol.~6, art.~1298330, 2023, doi: 10.3389/fbloc.2023.1298330.

	\bibitem{mckenzie2017almp} D. McKenzie, ``How effective are active labor market policies in developing countries? A critical review of recent evidence,'' \emph{The World Bank Research Observer}, vol.~32, no.~2, pp.~127--154, 2017, doi: 10.1093/wbro/lkx001.

\end{thebibliography}
\vspace{12pt}

\end{document}

